---
title: "Coral_Reef_Case_Study"
author: "George Ni"
date: "10/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 3


```{r}
require(deSolve)

#adding comments
state <- c(N=50) #At N=10 and N=50, the population stays static, which could be eq? Starting at N-8, the population goes down to 0
times <- seq(0,100,by=0.1)

parameters <- c(b = 2.4,c=0.02,M=50,d=0.2)

sterile_insect <- function(t,state,parameters){
  with(as.list(c(state,parameters)),{
    dN <- b*N/(N+M) - d - c*N
    list(c(dN))   
  })}

out <- ode(y = state,times=times,func=sterile_insect,parms=parameters)

result<-vector("numeric", 60)
for (i in 1:60) {
  dN<-dN <- 2.4*i/(i+50) - 0.2 - 0.02*i
  result[i]<-dN
}

N<-seq(from=1, to=60, by=1)
plot(x=N, y=result, lines(x = N,y = result), ylab="dN/dt", xlab="N", cex=0.1)
points(x=10, y=0, pch=1, cex=2)
points(x=50, y=0, pch=16, cex=2)
abline(h=0)

```


```{r}
require(deSolve)

# Initial values
state <- c(M=10,P=10,T=10,C=10)
times <- seq(0,100,by=0.1)



# Parameters
parameters <- c(a = 0.1,g=0.8,r=1.0,d=0.44,s=0.49,K=10)


# Model
coral_model <- function(t,state,parameters){
  with(as.list(c(state,parameters)),{
    dM <- (a*M*C) -((P*M)/(M + T))+ (g*M*T)
    dC <- (r*T*C)-(d*C)-(a*M*C)
    dT <- ((P*M)/(M + T))-(g*M*T)-(r*T*C)+(d*C)
    dP <- (s*P)*(1-(P/K))
    list(c(dM,dC,dT,dP))   
  })}



# Solve model and plot results
out <- ode(y = state,times=times,func=coral_model,parms=parameters)
par(mfrow=c(1,1))
plot(out,ylab='N',xlab='time',las=1)