---
title: "Problem Set 3"
author: "Liza Morse"
date: "10/2/2020"
output:
  pdf_document: default
  html_document: default
---

1) What species (and location if available) did you choose? Why did you choose the species?

We chose the peregrine falcon population studied in Capetown, South Africa. As both team members were familiar with peregrine falcons and their conservation history in the U.S, they thought it would be interesting to study effect of introducing nestboxes to peregrine falcon and determining their effect on population status. 

2) What question do you want to answer about this population (e.g. population status, best management strategies)?

Are nestboxes an effective strategy to increase Peregrine Falcon population in the Cape region of South Africa? If nest boxes were an effective strategy, we would expect to see a higher proportion of Subadults and Breeders age classes in the stable distribution of the "nestbox" population as compared to the natural breeding population.    

3) Calculate eigenvalue, stable age distribution, elasticity, and sensitivity. What does this tell you about the population?

```{r}
library(popbio)
library(popdemo)

load("COMADRE_v.4.20.9.0 (1).rdata")

# Calculations for Nestbox Data:

comadre$matrixClass[[835]]$MatrixClassAuthor

comadre$metadata$CommonName
comadre$mat[[835]]$matA #835 is 1997-2010 for Nestbox data

lambdaVal_nestbox <-eigs(comadre$mat[[835]]$matA, what="lambda") #dominant eigenvalue
#   -0.14969159346691008 for other lambda value
rightVal_nestbox <-eigs(comadre$mat[[835]]$matA, what="ss") #eigen vector (aka stable age distribution)
rightVal_nestbox

sense<-sensitivity(comadre$mat[[835]]$matA)
elast<-elasticity(comadre$mat[[835]]$matA)
#Sensitivity-
```

```{r}
# Calculations for Unmanipulated Data:

comadre$mat[[849]]$matA #849 is 1997-2010 for unmanipulated data data

lambdaVal_unmanipulated <-eigs(comadre$mat[[849]]$matA, what="lambda") #eigenvalue

rightVal_unmanipulated <-eigs(comadre$mat[[849]]$matA, what="ss") #eigen vector (aka stable age distribution)
rightVal_unmanipulated

sense2<-sensitivity(comadre$mat[[849]]$matA)
elast2<-elasticity(comadre$mat[[849]]$matA)
```

4) Using the calculations in part (c), or additional calculations, address the question you proposed in part (b).

Eigenvalues for nestboxes and unmanipulated populations were 1.03127, and 0.9368. The stable age distributions, or right eigenvector, are shown below: 

```{r}
rightVal_nestbox
rightVal_unmanipulated
```
One thing to note is that the stable age distribution for the nestboxes' subadult 3 is 0, while it is 0.0055 
The sensitivities: 

```{r}
sense<-sensitivity(comadre$mat[[835]]$matA)
sense2<-sensitivity(comadre$mat[[849]]$matA)
```

And the elasticities:

```{r}
elast<-elasticity(comadre$mat[[835]]$matA)
elast2<-elasticity(comadre$mat[[849]]$matA)
```

