---
title: "Wright_Fisher_Model_Ni"
author: "George Ni"
date: "11/6/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Wright-Fisher model with genetic drift and mutation
# Created by Easton R. White
# Last edited: 4-Nov-2020

```{r}
N=500      # population size 
p_ini= 0.2   # the initial frequency of the A allele
u = 0.1      # mutation rate from a to A
max.time=100 # time to run simulations


sample.population <- function(p,N){
  next_pop = vector('character',length = N)
  for (i in 1:N){
    new_allele = sample(x = c('A','a'),size = 1,prob=c(p,1-p))
    if (new_allele=='a' & rbinom(1,1,prob = u)){new_allele='A'} # Optional line to include mutation
    next_pop[i] = as.character(new_allele)
  }
  return(sum(next_pop=='A')/N) # returns the value for p
}



p = vector(mode = 'numeric',length = max.time)
p[1]=p_ini

for (t in 1:(max.time-1)){
  p[t+1] = sample.population(p[t],N)
}

plot(1:max.time,p,ylim=c(0,1),type='l',las=1,ylab='Freq(A)',xlab='Time')
```

## Part 1 
# 1. What is the effect of population size (N) on the probability of extinction of A allele?

Increasing population size (N) decreases the probability of extinction of the A allele. 

# 2. How does the initial frequency of the A allele affect the probability it will reach fixation?

Increasing the initial frequency of the A allele dramatically increases the probability that it will reach fixation, and decreases in initial frequency increase the probability of reaching extinction. 


# 3. Let's now study the combined effects of mutation and genetic drift. Modify the above code to include a probability, u, that if a small "a" allele is chosen, it will mutate to be a big A allele. How does this affect our findings in questions 1 and 2?

By including the probability u that small "a" allele might mutate to be a big A allele,increases in that probability speed up the rate at which a particular allele reaches fixation. 




## PART 2: Quantitative Genetics with Reed deer on island of Jersey

## Key stats
## Population established there 6000 years ago
## Change in mean body weight from 220kg to 36kg
## 5 year generation time, heritability of 0.5



## 1. What is the mean change in body weight per generation?

Mean change in body weight per generation. 6000 years/5 years/gen = 1200 generations
220-36=184 kg change in mean body weight. Assuming equal reduction in body weight over time, we see a 0.1533 change in body weight per generation. 

## 2. What is the mean change in body weight caused by selection (estimate S)?
 TO calculate the effect of mean change caused by selection, we utilize the breeder's equation, $R=h^2*S$ . Knowing R, which is the mean change per generation, we divide that by our h2, or heritability, to estimate S. $R/h^2=S$. $0.1533/0.5=S$ $S=0.3066$

## 3. What assumptions are we making in these calculations?
We're assuming that there are equal reductions in body weight over time. Also, we are assuming that a lot of variables are static, such as there being no mutations, that all individuals are breeding. 
  

